### 2024-04-18 paper/paper.qmd
Updated

---
title: "Beyond Numbers: The Evolution of Mental Health in Canada Over Two Decades"
subtitle: "My subtitle if needed"
author: 
  - Zijun Meng
thanks: "Code and data are available at: https://github.com/FrankMengZJ/canada_mental_health/"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Mental health has increasingly become a focal point of public health discussions globally, and Canada is no exception. Over the last two decades, shifts in societal attitudes, healthcare policies, and demographic dynamics have contributed significantly to changes in the mental health landscape. The Canadian Mental Health Association highlights a growing crisis, with one in five Canadians experiencing a mental health problem or illness each year, underscoring the critical need for comprehensive research into this area (CAMH, 2023). *Don't forget to cite*

This paper seeks to examine the evolution of mental health in Canada over the past twenty years, focusing on how sociodemographic factors such as age, gender, and geographic location influence mental health trends. Despite extensive studies in mental health, there remains a significant gap in longitudinal analyses that connect these shifts to broader sociodemographic changes. This study aims to fill this gap by leveraging historical data to provide a detailed overview of the trends and factors that have shaped the mental health landscape in Canada during this period.

In conducting this research, we utilized data from various national databases, analyzing changes in mental health indicators across different demographics from 2002 to 2022. This approach not only provides a comprehensive view of the historical trends but also highlights the critical drivers behind these changes. Although the findings are pending detailed analysis, the research is poised to offer insights into the effectiveness of past interventions and the need for tailored future strategies in mental health services.

Understanding these trends is crucial for policymakers, healthcare providers, and community leaders to develop targeted interventions that address the unique challenges faced by diverse groups within Canada. The structure of the paper is as follows: Section 1 introduces the study’s background and objectives; @sec-data describes the methodology and data sources used; Section 3 presents the preliminary results; followed by a discussion in Section 4 of the implications of these findings for current and future mental health policies.

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-04-18 scripts/02-data_cleaning.R

@@ -11,11 +11,12 @@
 library(tidyverse)
 library(arrow)
 #### Clean data ####
-mental_health_indicators <- read_csv("data/raw_data/mental_health_indicators.csv")
-province <- read_csv("data/raw_data/province.csv")
-sociodemographic_characteristics <- read_csv("data/raw_data/sociodemographic_characteristics.csv")
+mental_health_indicators <- read_csv("data/raw_data/13100465.csv")
+province <- read_csv("data/raw_data/45100079.csv")
+sociodemographic_characteristics <- read_csv("data/raw_data/45100080.csv")
 
-
+province
+distinct(province,GEO)
 
 cleaned_data <- 
   filter(mental_health_indicators, Characteristics == 'Number of persons'| Characteristics == 'Percent' )

### 2024-04-18 scripts/03-test_data.R

@@ -16,3 +16,5 @@ library(arrow)
 mental_health_indicators <- read_parquet(file = "data/analysis_data/mental_health_indicators.parquet")
 province <- read_parquet(file = "data/analysis_data/province.parquet")
 sociodemographic_characteristics <- read_parquet(file = "data/analysis_data/sociodemographic_characteristics.parquet")
+
+province

### 2024-04-18 scripts/02-data_cleaning.R
Updated

@@ -14,6 +14,7 @@ library(arrow)
 mental_health_indicators <- read_csv("data/raw_data/13100465.csv")
 province <- read_csv("data/raw_data/45100079.csv")
 sociodemographic_characteristics <- read_csv("data/raw_data/45100080.csv")
+age <- read_csv("data/raw_data/13100096.csv")
 
 province
 distinct(province,GEO)
@@ -26,4 +27,4 @@ cleaned_data
 write_parquet(x=cleaned_data,sink="data/analysis_data/mental_health_indicators.parquet")
 write_parquet(x=province,sink="data/analysis_data/province.parquet")
 write_parquet(x=sociodemographic_characteristics,sink="data/analysis_data/sociodemographic_characteristics.parquet")
-
+write_parquet(x=age,sink="data/analysis_data/age.parquet")

### 2024-04-18 scripts/03-test_data.R

@@ -18,3 +18,6 @@ province <- read_parquet(file = "data/analysis_data/province.parquet")
 sociodemographic_characteristics <- read_parquet(file = "data/analysis_data/sociodemographic_characteristics.parquet")
 
 province
+
+print(province, n=10)
+variable.names(province)

### 2024-04-18 paper/paper.qmd
Update

@@
 ---
-title: "Beyond Numbers: The Evolution of Mental Health in Canada Over Two Decades"
+title: "Navigating Through Crisis: The Influence of COVID-19 on Canada’s Mental Health Problem"
 subtitle: "My subtitle if needed"
@@
 library(tidyverse)
-library(palmerpenguins)
+library(arrow)
+library(here)
+library(knitr)
+age <- read_parquet(file = here("data/analysis_data/age.parquet"))
+mental_health_indicators <- read_parquet(file = here( "data/analysis_data/mental_health_indicators.parquet"))
+province <- read_parquet(file = here("data/analysis_data/province.parquet"))
+sc <- read_parquet(file = here("data/analysis_data/sociodemographic_characteristics.parquet"))
 ```
@@
 
-Mental health has increasingly become a focal point of public health discussions globally, and Canada is no exception.
-Over the last two decades, shifts in societal attitudes, healthcare policies, and demographic dynamics have contributed significantly to changes in the mental health landscape.
-The Canadian Mental Health Association highlights a growing crisis, with one in five Canadians experiencing a mental health problem or illness each year, underscoring the critical need for comprehensive research into this area (CAMH, 2023).
-*Don't forget to cite*
+The global COVID-19 pandemic has had a profound impact on public health, particularly in the realm of mental health.
+In Canada, the challenges brought on by the pandemic have intensified existing mental health issues and revealed significant vulnerabilities within the healthcare system (Government of Canada, 2021) *cite this*.
+This paper explores the effects of the COVID-19 pandemic on mental health trends in Canada, with a particular focus on the shifts in prevalence and the disparities in healthcare access observed across different sociodemographic groups during this period.
 
-This paper seeks to examine the evolution of mental health in Canada over the past twenty years, focusing on how sociodemographic factors such as age, gender, and geographic location influence mental health trends.
-Despite extensive studies in mental health, there remains a significant gap in longitudinal analyses that connect these shifts to broader sociodemographic changes.
-This study aims to fill this gap by leveraging historical data to provide a detailed overview of the trends and factors that have shaped the mental health landscape in Canada during this period.
+Research on the long-term mental health trends during the COVID-19 pandemic is sparse, particularly studies that provide a comprehensive analysis across various sociodemographic groups.
+Most existing studies have concentrated on immediate or short-term impacts within narrowly defined populations.
+This study aims to bridge this gap by conducting a detailed examination of how factors such as age, gender, and geographic location have influenced mental health outcomes across the broader Canadian population during the pandemic (CAMH, 2020) *cite this*.
 
-In conducting this research, we utilized data from various national databases, analyzing changes in mental health indicators across different demographics from 2002 to 2022.
-This approach not only provides a comprehensive view of the historical trends but also highlights the critical drivers behind these changes.
-Although the findings are pending detailed analysis, the research is poised to offer insights into the effectiveness of past interventions and the need for tailored future strategies in mental health services.
+The methodology of this paper includes a comparative analysis of data from national health surveys, focusing on the period from 2019 to the latest available data in 2021.
+This approach allows us to observe both pre-pandemic and during-pandemic mental health indicators, identifying significant trends and changes.
+Preliminary observations suggest that there are significant disparities in the impact of the pandemic on mental health, which vary widely among different demographic groups.
 
-Understanding these trends is crucial for policymakers, healthcare providers, and community leaders to develop targeted interventions that address the unique challenges faced by diverse groups within Canada.
-The structure of the paper is as follows: Section 1 introduces the study’s background and objectives; @sec-data describes the methodology and data sources used; Section 3 presents the preliminary results; followed by a discussion in Section 4 of the implications of these findings for current and future mental health policies.
+This research is critical not only for academic purposes but also for policymakers and healthcare providers planning interventions for ongoing and future public health crises.
+The structure of the paper is as follows: Section 1 introduces the background and literature relevant to the study; Section 2 outlines the methodology and data sources; Section 3 presents the analysis of the findings; Section 4 discusses the implications of these findings for mental health policy and practice in Canada.
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+There are four datasets used for this paper, including the 'Mental health indicator' dataset which contains data related to mental health in 2002 and 2012, 'Perceived mental health, by age group' dataset which data between 2015 and 2022, 'Perceived mental health, by gender and other selected sociodemographic characteristics' and 'Perceived mental health, by gender and province' datasets both contain data between 2021 and 2023, but have different variables.
 
+## Mental Health Indicator Dataset
+
+The mental health indicator dataset 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: tbl-1
+#| fig-cap: Mental Health Indicator Dataset 
 #| echo: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+kable(head(mental_health_indicators,5))
 ```

### 2024-04-18 scripts/03-test_data.R

@@ -13,6 +13,7 @@ library(tidyverse)
 library(arrow)
 
 #### Test data ####
+age <- read_parquet(file = "data/analysis_data/age.parquet")
 mental_health_indicators <- read_parquet(file = "data/analysis_data/mental_health_indicators.parquet")
 province <- read_parquet(file = "data/analysis_data/province.parquet")
 sociodemographic_characteristics <- read_parquet(file = "data/analysis_data/sociodemographic_characteristics.parquet")

### 2024-04-18 scripts/02-data_cleaning.R
Update

@@ -16,13 +16,14 @@ province <- read_csv("data/raw_data/45100079.csv")
 sociodemographic_characteristics <- read_csv("data/raw_data/45100080.csv")
 age <- read_csv("data/raw_data/13100096.csv")
 
-province
-distinct(province,GEO)
 
 cleaned_data <- 
-  filter(mental_health_indicators, Characteristics == 'Number of persons'| Characteristics == 'Percent' )
-cleaned_data
+  filter(mental_health_indicators, Characteristics == 'Percent'|Characteristics == 'Number of persons' )
 
+cleaned_data <-
+  cleaned_data |>
+  rename(Age='Age group') |>
+  select(REF_DATE, GEO, Age, Sex, Indicators,Characteristics, VALUE)
 #### Save data ####
 write_parquet(x=cleaned_data,sink="data/analysis_data/mental_health_indicators.parquet")
 write_parquet(x=province,sink="data/analysis_data/province.parquet")

### 2024-04-18 scripts/02-data_cleaning.R
Updated data cleaning

@@ -13,19 +13,42 @@ library(arrow)
 #### Clean data ####
 mental_health_indicators <- read_csv("data/raw_data/13100465.csv")
 province <- read_csv("data/raw_data/45100079.csv")
-sociodemographic_characteristics <- read_csv("data/raw_data/45100080.csv")
+sc <- read_csv("data/raw_data/45100080.csv")
 age <- read_csv("data/raw_data/13100096.csv")
 
-
-cleaned_data <- 
+# Clean 'Mental health indicators' dataset
+mental_health_indicators <- 
   filter(mental_health_indicators, Characteristics == 'Percent'|Characteristics == 'Number of persons' )
 
-cleaned_data <-
-  cleaned_data |>
+mental_health_indicators <-
+  mental_health_indicators |>
   rename(Age='Age group') |>
   select(REF_DATE, GEO, Age, Sex, Indicators,Characteristics, VALUE)
+
+# Clean 'Perceived mental health, by age group' dataset
+age <-
+  age |>
+  filter(Characteristics == 'Percent') |>
+  rename(Age='Age group') |>
+  select(REF_DATE, GEO, Age, Sex, Indicators, VALUE)
+
+# Clean 'Perceived mental health, by gender and other selected sociodemographic characteristics' dataset
+
+province <-
+  province |>
+  filter(Statistics == 'Percentage of persons') |>
+  select(REF_DATE, GEO, Gender, Indicators, VALUE)
+
+# Clean 'Perceived mental health, by gender and province' dataset
+sc <- 
+  sc |>
+  rename(Sociodemographic_characteristics = 'Sociodemographic characteristics' )|>
+  filter(Statistics == 'Percentage of persons') |>
+  select(REF_DATE, GEO, Gender, Sociodemographic_characteristics, Indicators, VALUE)
+  
+
 #### Save data ####
-write_parquet(x=cleaned_data,sink="data/analysis_data/mental_health_indicators.parquet")
+write_parquet(x=mental_health_indicators,sink="data/analysis_data/mental_health_indicators.parquet")
 write_parquet(x=province,sink="data/analysis_data/province.parquet")
-write_parquet(x=sociodemographic_characteristics,sink="data/analysis_data/sociodemographic_characteristics.parquet")
+write_parquet(x=sc,sink="data/analysis_data/sociodemographic_characteristics.parquet")
 write_parquet(x=age,sink="data/analysis_data/age.parquet")

### 2024-04-18 scripts/03-test_data.R

@@ -12,13 +12,12 @@
 library(tidyverse)
 library(arrow)
 
-#### Test data ####
 age <- read_parquet(file = "data/analysis_data/age.parquet")
 mental_health_indicators <- read_parquet(file = "data/analysis_data/mental_health_indicators.parquet")
 province <- read_parquet(file = "data/analysis_data/province.parquet")
 sociodemographic_characteristics <- read_parquet(file = "data/analysis_data/sociodemographic_characteristics.parquet")
 
-province
+#### Test data ####
 
-print(province, n=10)
-variable.names(province)
+age
+variable.names(age)

### 2024-04-18 scripts/02-data_cleaning.R
Updated data cleaning

@@ -18,12 +18,12 @@ age <- read_csv("data/raw_data/13100096.csv")
 
 # Clean 'Mental health indicators' dataset
 mental_health_indicators <- 
-  filter(mental_health_indicators, Characteristics == 'Percent'|Characteristics == 'Number of persons' )
+  filter(mental_health_indicators, Characteristics == 'Percent' )
 
 mental_health_indicators <-
   mental_health_indicators |>
   rename(Age='Age group') |>
-  select(REF_DATE, GEO, Age, Sex, Indicators,Characteristics, VALUE)
+  select(REF_DATE, GEO, Age, Sex, Indicators, VALUE)
 
 # Clean 'Perceived mental health, by age group' dataset
 age <-
@@ -52,3 +52,8 @@ write_parquet(x=mental_health_indicators,sink="data/analysis_data/mental_health_
 write_parquet(x=province,sink="data/analysis_data/province.parquet")
 write_parquet(x=sc,sink="data/analysis_data/sociodemographic_characteristics.parquet")
 write_parquet(x=age,sink="data/analysis_data/age.parquet")
+
+write_csv(mental_health_indicators,"mental_health_indicators.csv")
+write_csv(province,"province.csv")
+write_csv(sc,"sc.csv")
+write_csv(age,"age.csv")

### 2024-04-18 paper/paper.qmd
Update
Updated data cleaning
Updated

@@
 
-The mental health indicator dataset 
+The mental health indicator dataset records 
 ```{r}
@@
 
-Talk more about it.
+# Data {#sec-data}
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+## Overview of Datasets
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
+The research utilizes four pivotal datasets sourced, each providing distinct insights into the mental health situation of Canada.
+These datasets were chosen due to their comprehensive coverage, relevance to the research questions, and the credibility of the source.
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+### Datasets Description
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
+#### Mental Health Indicator Dataset (`mental_health_indicators.csv`)
+This dataset covers the years 2002 and 2012 and includes variables such as the prevalence of specific mental health conditions, access to mental health services, and outcomes of mental health interventions.
+The data provides a snapshot of mental health status at two points a decade apart, enabling an analysis of trends and changes over time.
 
-Talk way more about it.
+Source: [Statistics Canada - Mental Health Indicator Dataset](https://doi.org/10.25318/4510008001-eng)
+
+#### 2. Perceived Mental Health, by Age Group (`age.csv`)
+Spanning from 2015 to 2022, this dataset details the perceived mental health statuses across different age cohorts.
+Variables include age-specific perceptions and reported mental health conditions, allowing for an age-wise analysis pertinent to assessing generational impacts on mental health, especially pre- and post-COVID-19.
+
+Source: [Statistics Canada - Perceived Mental Health by Age Group](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=3226)
+
+#### 3. Perceived Mental Health, by Gender and Sociodemographic Characteristics (`sc.csv`)
+Capturing data from 2021 to 2023, this dataset includes detailed breakdowns by gender along with other sociodemographic factors such as socioeconomic status, education, and employment.
+These variables are crucial for understanding the intersectional impacts of various social determinants on mental health.
+
+Source: [Statistics Canada - Survey on Mental Health by Sociodemographic Characteristics](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5354)
+
+#### 4. Perceived Mental Health, by Gender and Province (`province.csv`)
+Also covering the period from 2021 to 2023, this dataset provides a geographic breakdown of mental health perceptions across Canadian provinces, categorized by gender.
+It enables a regional analysis which is essential for identifying area-specific challenges and needs in mental health services.
+
+Source: [Open Canada - Perceived Mental Health by Province](https://open.canada.ca/data/en/dataset/a9863f45-6a1a-4277-ae0f-2d9cb61d413a)
+
+## Data Selection Rationale
+
+These datasets were selected over other potential sources due to their direct relevance to the research objectives, encompassing both specific and broad factors influencing mental health.
+They provide both longitudinal data and recent snapshots that are vital for understanding the ongoing impacts of the COVID-19 pandemic on mental health across diverse populations.
+The chosen datasets offer the most current and comprehensive data available from a reliable national source, ensuring the robustness and relevance of the analysis.
+
+## Data Exploration and Summary Statistics
+
+*For each dataset, a series of descriptive statistics and graphical analyses will be presented.* These will include:
+- **Distribution graphs** for each variable to understand the data spread and central tendencies.
+- **Cross-tabulations and correlation matrices** to explore relationships between variables.
+- **Summary statistics tables** summarizing key metrics such as mean, median, mode, range, and standard deviation for each variable.
+
+The detailed graphs and tables will be included in the main text or appendices, depending on the level of detail.
+Each figure and table will be accompanied by a discussion interpreting the findings and their implications for understanding mental health trends in Canada.
+
+## Data Preparation
+
+The datasets underwent essential cleaning steps including handling missing values, correcting any inconsistencies in data reporting, and standardizing variable formats across datasets to ensure compatibility for analysis.
+Variables of interest were constructed based on raw data to better align with the research questions, such as aggregating individual responses into broader categories relevant to the study.
+
+The thorough examination and explanation of variables, alongside the visual presentation of the data, ensure a comprehensive understanding of the datasets and support the reliability and validity of the subsequent analysis.
 

### 2024-04-18 paper/paper.qmd
Updated

--- 2024-04-18 paper/paper.qmd
@@
 
-The global COVID-19 pandemic has had a profound impact on public health, particularly in the realm of mental health.
-In Canada, the challenges brought on by the pandemic have intensified existing mental health issues and revealed significant vulnerabilities within the healthcare system (Government of Canada, 2021) *cite this*.
-This paper explores the effects of the COVID-19 pandemic on mental health trends in Canada, with a particular focus on the shifts in prevalence and the disparities in healthcare access observed across different sociodemographic groups during this period.
+The COVID-19 pandemic has spotlighted mental health as a critical component of public health, magnifying existing challenges and uncovering gaps within healthcare systems globally.
+In Canada, the impact of the pandemic on mental health has been profound, exacerbating disparities and highlighting the urgency of effective mental health care.
+The Centre for Addiction and Mental Health (CAMH) has identified a significant increase in mental health problems among Canadians, noting that the pandemic has led to unprecedented levels of anxiety and depression (CAMH, 2020).
+Additionally, the Public Health Agency of Canada reports on the elevated symptoms of anxiety and depression during the pandemic, further underscoring the need for comprehensive data analysis on this issue (Public Health Agency of Canada, 2021).
 
-Research on the long-term mental health trends during the COVID-19 pandemic is sparse, particularly studies that provide a comprehensive analysis across various sociodemographic groups.
-Most existing studies have concentrated on immediate or short-term impacts within narrowly defined populations.
-This study aims to bridge this gap by conducting a detailed examination of how factors such as age, gender, and geographic location have influenced mental health outcomes across the broader Canadian population during the pandemic (CAMH, 2020) *cite this*.
+This paper aims to provide a detailed analysis of the changing situation of mental health in Canada by using four datasets spanning from 2002 to 2023.
+The focus is on perceived mental health statuses across various demographic segments, including age, gender, and province.
+The primary objective is to understand how perceptions of mental health have evolved over the past two decades and how these perceptions have been impacted by the COVID-19 pandemic.
 
-The methodology of this paper includes a comparative analysis of data from national health surveys, focusing on the period from 2019 to the latest available data in 2021.
-This approach allows us to observe both pre-pandemic and during-pandemic mental health indicators, identifying significant trends and changes.
-Preliminary observations suggest that there are significant disparities in the impact of the pandemic on mental health, which vary widely among different demographic groups.
+The study addresses the gap in longitudinal mental health research, particularly in the analysis of sociodemographic influences during a major global crisis.
+Through the comparison of mental health indicators before and during the pandemic, this research provides insights into the shifts in mental health perceptions and the effectiveness of the responses to these changes.
+Initial findings suggest notable disparities in mental health impacts across different demographic groups, indicating the need for targeted interventions.
 
-This research is critical not only for academic purposes but also for policymakers and healthcare providers planning interventions for ongoing and future public health crises.
-The structure of the paper is as follows: Section 1 introduces the background and literature relevant to the study; Section 2 outlines the methodology and data sources; Section 3 presents the analysis of the findings; Section 4 discusses the implications of these findings for mental health policy and practice in Canada.
+The significance of this research lies in its contribution to the evidence base informing mental health policy and practice.
+By identifying the demographic factors associated with varying mental health outcomes, the study aids policymakers and healthcare providers in developing tailored strategies to address the unique challenges faced by diverse groups within Canada.
+*TBD*
 
@@
 
-Talk more about it.
+# Data {#sec-data}
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+This study utilizes four carefully curated datasets from Statistics Canada, each providing valuable insights into the mental health perceptions of Canadians.
+These datasets are pivotal for understanding the evolution of mental health across different demographics and geographical regions from 2002 to 2023.
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
+## Overview of Datasets
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+The data for this research come from national surveys conducted by Statistics Canada, specifically designed to monitor various aspects of mental health across the country.
+Each dataset targets different demographic variables, such as age, gender, and province, and covers distinct periods critical for analyzing long-term trends and the immediate effects of the COVID-19 pandemic.
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
+### Mental Health Indicators Dataset
+Data from 2002 and 2012 in the Mental Health Indicators Dataset were derived from national health databases, capturing key statistics on the prevalence of mental disorders and utilization of mental health services.
+The information is gathered through comprehensive health surveys and administrative records, providing robust measures of mental health status and service outcomes.
 
-Talk way more about it.
+### Perceived Mental Health by Age Group
+The Perceived Mental Health by Age Group dataset spans from 2015 to 2022 and utilizes a consistent methodological framework to assess self-reported mental health.
+The data collection process involves standardized questionnaires distributed to a representative sample of Canadian adults across various age groups.
+This approach ensures that the data reflect a broad spectrum of the population, excluding those in full-time residence, on Indigenous reserves, or non-permanent residents, thus focusing on a generalizable sample of the Canadian populace.
+
+### Perceived Mental Health by Gender and Sociodemographic Characteristics
+The data for 2021 to 2023 from this dataset is sourced from the Canadian Community Health Survey (CCHS), which meticulously gathers data on mental health perceptions influenced by gender and other sociodemographic factors such as income, education, and employment.
+This survey selects participants through a stratified sampling design aimed at capturing diverse sociodemographic landscapes across Canada, enhancing the dataset's relevance for analyzing intersectional impacts on mental health.
+
+## Detailed Measurement Discussion
+
+Statistics Canada employs rigorous methodologies to ensure data accuracy and reliability.
+For example, the survey instruments are tested extensively before deployment to ensure they accurately capture the intended data.
+Additionally, data collectors are trained to adhere to strict guidelines to maintain consistency and reduce potential biases during data collection.
+
+The sample for each dataset is designed to be representative of the Canadian population, stratified by key demographic variables to ensure comprehensive coverage across different regions and groups.
+This stratification helps in understanding specific sub-population dynamics within the broader national context.
+Data is then weighted to adjust for potential non-response biases and to represent the population accurately.
+
+Despite these rigorous methods, the reliability of self-reported data can sometimes be challenging to assess due to the subjective nature of mental health perceptions.
+Statistics Canada addresses this by implementing multiple stages of data verification and validation, including automated data checks and manual reviews by experts.
+Furthermore, unusual or outlier responses undergo additional scrutiny to confirm their validity.
 

### 2024-04-18 paper/paper.qmd
Updated data testing

@@
 
-## Detailed Measurement Discussion
+### Detailed Measurement Discussion
 
-Statistics Canada employs rigorous methodologies to ensure data accuracy and reliability.
-For example, the survey instruments are tested extensively before deployment to ensure they accurately capture the intended data.
-Additionally, data collectors are trained to adhere to strict guidelines to maintain consistency and reduce potential biases during data collection.
+Both the CCHS and other Statistics Canada surveys employ rigorous data verification and validation processes.
+Quality control measures include logical checks during data collection and detailed follow-ups for non-responses.
+Post-stratification weighting adjusts for any discrepancies due to non-response rates, and seasonal adjustments are made to account for the timing of data collection.
+These methodologies ensure that the data are robust, reliable, and accurately reflect the Canadian population’s health status.
 
-The sample for each dataset is designed to be representative of the Canadian population, stratified by key demographic variables to ensure comprehensive coverage across different regions and groups.
-This stratification helps in understanding specific sub-population dynamics within the broader national context.
-Data is then weighted to adjust for potential non-response biases and to represent the population accurately.
+The CCHS's use of health-specific, standardized questionnaires is designed to capture nuanced information on health determinants and outcomes, which is critical for public health research and intervention planning.
+Conversely, Statistics Canada’s broader surveys provide comprehensive data across a wider range of topics, offering valuable insights into the interplay between health and other societal factors.
 
-Despite these rigorous methods, the reliability of self-reported data can sometimes be challenging to assess due to the subjective nature of mental health perceptions.
-Statistics Canada addresses this by implementing multiple stages of data verification and validation, including automated data checks and manual reviews by experts.
-Furthermore, unusual or outlier responses undergo additional scrutiny to confirm their validity.
+### Comparison of Data Collection Specifics
 
-# Model
+CCHS is highly adaptable with its questionnaire content, allowing for the inclusion of regional health issues of specific interest, which is crucial for targeted public health interventions.
+This adaptability in data collection, combined with its detailed focus on health, makes CCHS uniquely suited for studies like ours that require specific health-related insights.
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-```{=tex}
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-```
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+In summary, the methodologies used by the Canadian Community Health Survey and other Statistics Canada surveys are both robust and tailored to meet different informational needs.
+Understanding these methodologies helps in accurately interpreting the data collected on mental health perceptions, ensuring that research findings are based on reliable and representative data.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+Our results are summarized in .
 
@@
 
-library(rstanarm)
+temp <- filter(age, Indicators == "Perceived mental health, fair or poor" & Age == "Total, 12 years and over" & Sex == "Both sexes" & GEO == "Canada (excluding territories)")
+ggplot(temp, aes(x=REF_DATE,y=VALUE))+
+  geom_line()+
+  labs(x = "Year", y = "Percentage", title = "Perceived mental health, fair or poor, all people over 12 years old in Canada") +
+  theme_minimal()
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
 ```
@@
 
-\appendix
-
-# Appendix {.unnumbered}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
-\newpage
-
 # References

### 2024-04-18 scripts/03-test_data.R

@@ -20,4 +20,14 @@ sociodemographic_characteristics <- read_parquet(file = "data/analysis_data/soci
 #### Test data ####
 
 age
-variable.names(age)
+distinct(age,REF_DATE)
+distinct(age, Indicators)
+
+filter(mental_health_indicators,Indicators == "Perceived mental health, fair or poor", GEO == "Canada", Age == "Total, 15 years and over", Sex == "Both sexes")
+
+filter(age, Indicators == "Perceived mental health, fair or poor" & Age == "Total, 12 years and over" & Sex == "Both sexes" & GEO == "Canada (excluding territories)")
+
+filter(province, GEO == "Canada (excluding territories)", Gender == "Total, all persons", Indicators == "Fair or poor perceived mental health")
+
+filter(sociodemographic_characteristics, GEO == "Canada (excluding territories)", Gender == "Total, all persons", Sociodemographic_characteristics == "Total, 15 years and over", Indicators == "Fair or poor perceived mental health")
+

### 2024-04-20 paper/paper.qmd
Updated results

@@
 
-There are four datasets used for this paper, including the 'Mental health indicator' dataset which contains data related to mental health in 2002 and 2012, 'Perceived mental health, by age group' dataset which data between 2015 and 2022, 'Perceived mental health, by gender and other selected sociodemographic characteristics' and 'Perceived mental health, by gender and province' datasets both contain data between 2021 and 2023, but have different variables.
-
-## Mental Health Indicator Dataset
-
-The mental health indicator dataset records 
-```{r}
-#| label: tbl-1
-#| fig-cap: Mental Health Indicator Dataset 
-#| echo: false
-
-kable(head(mental_health_indicators,5))
-```
-
-# Data {#sec-data}
-
-This study utilizes four carefully curated datasets from Statistics Canada, each providing valuable insights into the mental health perceptions of Canadians.
+This study uses four carefully curated datasets from Statistics Canada, each providing valuable insights into the mental health perceptions of Canadians.
 These datasets are pivotal for understanding the evolution of mental health across different demographics and geographical regions from 2002 to 2023.
@@
 The information is gathered through comprehensive health surveys and administrative records, providing robust measures of mental health status and service outcomes.
+
+@tbl-1 shows the first 5 rows in the dataset, REF_DATE stands for the time that the data was gathered, GEO denotes the geographical ,
+
+```{r}
+#| label: tbl-1
+#| tbl-cap: First 5 Rows in Mental Health Indicators Dataset 
+#| echo: false
+
+kable(head(mental_health_indicators,5))
+```
 
@@
 
+## Overview
+
 ```{r}
@@
 #| message: false
+#| label: fig-trend1
+#| fig-cap: Perceived mental health fair or poor in Canada, 2002-2022
 
-temp <- filter(age, Indicators == "Perceived mental health, fair or poor" & Age == "Total, 12 years and over" & Sex == "Both sexes" & GEO == "Canada (excluding territories)")
+mhi <- filter(mental_health_indicators, Indicators == "Perceived mental health, fair or poor" & Age == "Total, 15 years and over" & Sex == "Both sexes" & GEO == "Canada")
+
+ag <- filter(age, Indicators == "Perceived mental health, fair or poor" & Age == "Total, 12 years and over" & Sex == "Both sexes" & GEO == "Canada (excluding territories)")
+
+temp <-
+  rbind(select(mhi,REF_DATE,VALUE),select(ag,REF_DATE,VALUE))
+
 ggplot(temp, aes(x=REF_DATE,y=VALUE))+
   geom_line()+
-  labs(x = "Year", y = "Percentage", title = "Perceived mental health, fair or poor, all people over 12 years old in Canada") +
-  theme_minimal()
+  geom_point()+
+  labs(x = "Year", y = "Percentage") +
+  theme_linedraw()
 
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
+#| message: false
+#| label: fig-trend2
+#| fig-cap: Perceived mental health fair or poor in Canada
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+temp <- filter(sc, Indicators == "Fair or poor perceived mental health" & Sociodemographic_characteristics  == "Total, 15 years and over" & Gender == "Total, all persons" & GEO == "Canada (excluding territories)")
+
+ggplot(temp, aes(x=REF_DATE,y=VALUE))+
+  geom_line(aes(group=1))+
+  geom_line()+
+  geom_point()+
+  labs(x = "Year", y = "Percentage") +
+  theme_linedraw()
+
+```
+## By Province
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-trend3
+#| fig-cap: Perceived mental health fair or poor in Canada
+
+temp <- filter(age, Indicators == "Perceived mental health, fair or poor" & Sex == "Both sexes", Age =="Total, 12 years and over")
+
+temp <-
+  temp|>
+  select(REF_DATE,GEO,VALUE)
+
+temp1 <-
+  group_by(temp, GEO) |>
+  summarise(variance = var(VALUE))
+
+ggplot(temp, aes(x= REF_DATE, y= VALUE, group = GEO, color = GEO))+
+  geom_line()+
+  theme_linedraw()
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-trend4
+#| fig-cap: Perceived mental health fair or poor in Canada
+
+temp <- filter(province, Indicators == "Fair or poor perceived mental health" & Gender == "Total, all persons")
+
+temp <-
+  temp|>
+  select(REF_DATE,GEO,VALUE)
+
+temp1 <-
+  group_by(temp, GEO) |>
+  summarise(variance = var(VALUE))
+
+ggplot(temp, aes(x= REF_DATE, y= VALUE, group = GEO, color = GEO))+
+  geom_line()+
+  theme_linedraw()
+
+```
+
+## By Gender
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-trend4
+#| fig-cap: Perceived mental health fair or poor in Canada
+
+temp <- filter(age, Indicators == "Perceived mental health, fair or poor" & GEO == "Canada (excluding territories)"& Age == "Total, 12 years and over")
+
+temp1 <-
+  group_by(temp, Sex)
+
+colors <- c("Males" = "blue", "Females" = "green4","Both sexes" = "red")
+
+ggplot(temp, aes(x= REF_DATE, y= VALUE, group = Sex, color = Sex))+
+  geom_line()+
+  geom_point()+
+  scale_color_manual(values = colors) +
+  theme_linedraw()
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-trend2
+#| fig-cap: Perceived mental health fair or poor in Canada
+
+temp <- filter(sc, Indicators == "Fair or poor perceived mental health" & Sociodemographic_characteristics  == "Total, 15 years and over" & GEO == "Canada (excluding territories)")
+
+colors <- c("Men" = "blue", "Women" = "green4","Total, all persons" = "red")
+
+ggplot(temp, aes(x=REF_DATE,y=VALUE, group = Gender, color = Gender))+
+  geom_line()+
+  geom_point()+
+  labs(x = "Year", y = "Percentage") +
+  scale_color_manual(values = colors) +
+  theme_linedraw()
+
+```
+## By Age
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-trend4
+#| fig-cap: Perceived mental health fair or poor in Canada
+
+temp <- filter(age, Indicators == "Perceived mental health, fair or poor" & GEO == "Canada (excluding territories)" & Sex == "Both sexes")
+
+ggplot(temp, aes(x= REF_DATE, y= VALUE, group = Age, color = Age))+
+  geom_line()+
+  geom_point()+
+  theme_linedraw()
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-trend2
+#| fig-cap: Perceived mental health fair or poor in Canada
+
+temp <- filter(sc, Indicators == "Fair or poor perceived mental health" & Gender  == "Total, all persons" & GEO == "Canada (excluding territories)" & grepl('years',Sociodemographic_characteristics)& !grepl("Immigrants",Sociodemographic_characteristics))
+
+ggplot(temp, aes(x=REF_DATE,y=VALUE, group = Sociodemographic_characteristics , color = Sociodemographic_characteristics ))+
+  geom_line()+
+  geom_point()+
+  labs(x = "Year", y = "Percentage") +
+  theme_linedraw()
+
+```
+
+## Others
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-trend2
+#| fig-cap: Perceived mental health fair or poor in Canada
+
+temp <- filter(sc, Indicators == "Fair or poor perceived mental health")
+
+temp <- drop_na(temp, VALUE)
+
+temp1 <- 
+  group_by(temp, GEO, Gender, Sociodemographic_characteristics) |>
+  summarise(avg = mean(VALUE))
+
+print(temp1[order(-temp1$avg),],n=10)
+
 ```

### 2024-04-22 paper/paper.qmd
Updated results

@@
 
-Our results are summarized in .
+## Methodological Considerations
+
+Prior to examining the trends, it's crucial to note that the datasets used for this analysis were collected using different methodologies.
+Annual data up to 2020 required a distinct approach from the quarterly data collected during the pandemic's peak.
+To preserve the accuracy of the analysis, the results are presented in separate graphs corresponding to these differing collection frequencies.
 
@@
 
+The results from the analyzed datasets reveal significant trends in the mental health landscape across Canada.
+Initially, the data show a gentle increase in reported mental health concerns from the early 2000s up until 2010.
+Post-2010, there is a noticeable dip, followed by a steady climb in mental health issues leading up to 2015.
+However, the most striking feature of this trend is the sharp and substantial escalation in mental health concerns starting from around 2020, which corresponds with the onset of the COVID-19 pandemic.
+
+@fig-trend1 plots the percentage of individuals reporting mental health concerns from 2005 to the present.
+From 2005 to around 2015, the trajectory exhibits minor fluctuations with a gradual increase.
+However, in the years following 2015, there is a distinct upward surge, with the percentages almost doubling by 2020 and continuing to rise steeply.
+This dramatic increase highlights the profound impact that the pandemic has had on mental health, as isolation, stress, economic uncertainty, and health-related fears took a toll on the population.
+
+@fig-trend2 provides a more granular view of the recent period, showing quarterly data from 2021 through to mid-2023.
+This period captures the ongoing effects of the pandemic.
+The graph illustrates considerable volatility, with sharp increases and decreases in the percentage of individuals reporting mental health issues.
+These oscillations suggest that mental health responses are highly sensitive to the evolving circumstances of the pandemic, such as waves of infection, public health measures, and societal changes.
+
+Together, these graphs paint a picture of a population that experienced a significant increase in mental health concerns during a global crisis.
+The data underscore the need for ongoing monitoring and support to address mental health at a community and national level.
+The results section will further dissect these trends by province, gender, age, and within special populations to understand the nuances of this mental health challenge.
 ```{r}
@@
 
+The results underscore the regional variations in mental health across Canadian provinces.
+In the graph illustrating data from 2016 to 2022, we see a common upward trend among all provinces, with a particularly steep increase following 2020, which aligns with the global onset of the COVID-19 pandemic.
+Prior to 2020, there were variations in the VALUE among provinces, but the gap between regions with the highest and lowest reported VALUE narrows as all regions trend upwards significantly in the subsequent years.
+
+Alberta, British Columbia, and Ontario exhibit some of the highest increases, suggesting a pronounced impact of the pandemic and its associated factors on mental health in these regions.
+However, it is notable that the trend is national, as even provinces like Prince Edward Island and Newfoundland and Labrador, which initially had some of the lowest reported VALUE, also show a sharp rise.
+
+The second graph, representing quarterly data from 2021 to 2023, provides a more nuanced picture with noticeable fluctuations within each province.
+These variations could be indicative of the changing conditions during the pandemic, such as waves of COVID-19 cases, differing provincial public health restrictions, and economic instability.
+While the Atlantic Region and the Prairies show a more volatile pattern, Quebec and Ontario demonstrate a more consistent increase over the observed quarters.
+
+Both graphs indicate that while the trend in mental health concerns is increasing across Canada, there are provincial specificities that suggest the impact of local factors.
+These could include economic conditions, the accessibility of mental health services, public health policies, and the social resilience of communities.
+The consistent rise across provinces post-2020, however, points to the overarching influence of the pandemic on mental health at a national level.
 ```{r}
@@
 #| message: false
-#| label: fig-trend3
+#| label: fig-province1
 #| fig-cap: Perceived mental health fair or poor in Canada
@@
 #| message: false
-#| label: fig-trend4
+#| label: fig-province2
 #| fig-cap: Perceived mental health fair or poor in Canada
@@
 
+The gender-based analysis reveals telling trends in the self-reported mental health status among Canadians, distinguishing between those who consider their mental health to be fair or poor.
+
+From 2016 to 2022, the first graph shows an upward trajectory for both males and females, with a notably sharper rise for females.
+By 2022, the percentage of females reporting their mental health as fair or poor substantially exceeds that of males, suggesting gender-specific impacts.
+The increase that coincides with the pandemic period may reflect the compounded stressors that women have faced, including potential increases in caregiving responsibilities, economic pressures, or other psychosocial factors exacerbated by the pandemic.
+
+In the quarterly data from 2021 to mid-2023, depicted in the second graph, the fluctuations are more pronounced, with both genders experiencing variability in their mental health reporting.
+The trend for women is characterized by higher peaks, indicating more pronounced shifts in mental health status as compared to men.
+This volatility could be influenced by the ebb and flow of pandemic-related stresses and public health measures.
+
+For both sexes combined, the overall trend is less erratic, pointing to the importance of disaggregating data by gender to fully capture the differential impacts on mental health.
+These findings highlight the significance of gender-responsive mental health resources and interventions, particularly in addressing the unique challenges that have emerged during the pandemic.
 ```{r}
@@
 ## By Age
+The first graph spans from 2016 to 2022 and illustrates the percentage of people within various age brackets reporting their mental health as fair or poor.
+A distinct trend is visible across all age groups, showing an increase in self-reported fair or poor mental health, with a marked acceleration from 2020 onwards, likely due to the COVID-19 pandemic.
+The oldest age group, those aged '65 years and over,' displays the most pronounced rise, indicating this demographic may have been particularly impacted during these years.
+Meanwhile, younger age groups also show a worrying increase, albeit less steep, suggesting widespread implications across generations.
+
+The second graph, which provides quarterly data from early 2021 to mid-2023, presents a more granular view, capturing the fluctuations within and between age groups during the pandemic.
+Notably, the '15 to 24 years' and '25 to 34 years' categories exhibit significant volatility, with peaks and valleys that could correlate with periods of pandemic waves, social restrictions, and economic impacts.
+The '65 years and over' group continues to report higher rates of fair or poor mental health, although with less fluctuation, possibly reflecting ongoing concerns about health risks and social isolation among seniors.
+
+These age-focused graphs suggest that mental health challenges have been a critical concern for all age groups, with particular implications for the most vulnerable populations.
+The rise in reported mental health issues among the youth and elderly underscores the need for age-specific mental health services and support systems tailored to their unique circumstances and stressors.
+The data call for continued monitoring and interventions to address the mental health repercussions of the pandemic across all age demographics.
+
 ```{r}
@@
 
+The examination of mental health within specific demographic segments reveals significant disparities.
+The most affected population, with 48.4%, are LGBTQ2+ women, followed by the total population of LGBTQ2+ individuals at 44.5%.
+LGBTQ2+ men report a lower percentage at 39.0%, indicating a gender-related variance within the LGBTQ2+ community.
+These figures highlight the substantial mental health challenges faced by LGBTQ2+ individuals, which may be attributed to factors such as discrimination, stigma, and lack of targeted support services.
+
+Southeast Asian women report a 39% rate of fair or poor mental health, indicating that cultural and possibly immigration-related factors might contribute to mental health concerns.
+This suggests the need for culturally sensitive mental health resources that are attuned to the unique experiences of minority groups.
+
+Young women aged 15 to 24 years report a rate of 36.3%, reflecting the particular vulnerabilities of this age group, which may include the pressures of social media, academic stress, and transitions into the workforce or higher education.
+
+Women with disabilities or long-term conditions report a 34.7% rate, with the total population of persons with disabilities reporting 32.4%.
+These figures underscore the intersectional nature of mental health, where disability and gender can compound the challenges faced by individuals, highlighting the critical need for accessible mental health services that accommodate physical and social barriers.
+
+Métis women, Indigenous women, and First Nations women report similar rates of around 30-31%, indicating the pressing mental health concerns among Indigenous populations.
+Historical context, ongoing systemic inequalities, and intergenerational trauma are likely contributing factors and underscore the importance of providing culturally competent care and support.
 ```{r}
@@
   group_by(temp, GEO, Gender, Sociodemographic_characteristics) |>
-  summarise(avg = mean(VALUE))
+  summarise(Average = mean(VALUE))
 
-print(temp1[order(-temp1$avg),],n=10)
+temp2 <- temp1[order(-temp1$Average),]
+kable(head(temp2,15))
 

### 2024-04-22 scripts/02-data_cleaning.R
Updated cleaning data

@@ -23,6 +23,7 @@ mental_health_indicators <-
 mental_health_indicators <-
   mental_health_indicators |>
   rename(Age='Age group') |>
+  filter(!is.na(VALUE)) |>
   select(REF_DATE, GEO, Age, Sex, Indicators, VALUE)
 
 # Clean 'Perceived mental health, by age group' dataset
@@ -43,7 +44,7 @@ province <-
 sc <- 
   sc |>
   rename(Sociodemographic_characteristics = 'Sociodemographic characteristics' )|>
-  filter(Statistics == 'Percentage of persons') |>
+  filter(Statistics == 'Percentage of persons' & !is.na(sc$VALUE)) |>
   select(REF_DATE, GEO, Gender, Sociodemographic_characteristics, Indicators, VALUE)
   
 
@@ -52,8 +53,3 @@ write_parquet(x=mental_health_indicators,sink="data/analysis_data/mental_health_
 write_parquet(x=province,sink="data/analysis_data/province.parquet")
 write_parquet(x=sc,sink="data/analysis_data/sociodemographic_characteristics.parquet")
 write_parquet(x=age,sink="data/analysis_data/age.parquet")
-
-write_csv(mental_health_indicators,"mental_health_indicators.csv")
-write_csv(province,"province.csv")
-write_csv(sc,"sc.csv")
-write_csv(age,"age.csv")

### 2024-04-22 scripts/03-test_data.R
Updated test data

@@ -19,15 +19,60 @@ sociodemographic_characteristics <- read_parquet(file = "data/analysis_data/soci
 
 #### Test data ####
 
-age
-distinct(age,REF_DATE)
+
+## Test age dataset
+
+age$Age |> min() >= 0
+
+age$REF_DATE |> min() == 2015
+age$REF_DATE |> max() == 2022
+
+distinct(age,Sex)
+
+distinct(age, GEO)
+
 distinct(age, Indicators)
 
-filter(mental_health_indicators,Indicators == "Perceived mental health, fair or poor", GEO == "Canada", Age == "Total, 15 years and over", Sex == "Both sexes")
 
-filter(age, Indicators == "Perceived mental health, fair or poor" & Age == "Total, 12 years and over" & Sex == "Both sexes" & GEO == "Canada (excluding territories)")
+## Test mental_health_indicators dataset
 
-filter(province, GEO == "Canada (excluding territories)", Gender == "Total, all persons", Indicators == "Fair or poor perceived mental health")
+mental_health_indicators$REF_DATE |> min() == 2002
+mental_health_indicators$REF_DATE |> max() == 2012
 
-filter(sociodemographic_characteristics, GEO == "Canada (excluding territories)", Gender == "Total, all persons", Sociodemographic_characteristics == "Total, 15 years and over", Indicators == "Fair or poor perceived mental health")
+mental_health_indicators$VALUE |> min() >= 0
 
+distinct(mental_health_indicators,Sex)
+
+distinct(mental_health_indicators,GEO)
+
+distinct(mental_health_indicators,Age)
+
+
+## Test province dataset
+
+as.Date(paste(province$REF_DATE,"-01", sep="")) |> min() == "2021-04-01"
+as.Date(paste(province$REF_DATE,"-01", sep="")) |> max() == "2023-07-01"
+
+province$VALUE |> min() >= 0
+
+distinct(province, Gender)
+
+distinct(province, GEO)
+
+distinct(province, Indicators)
+
+
+## Test sociodemographic_characteristics dataset
+
+as.Date(paste(sociodemographic_characteristics$REF_DATE,"-01", sep="")) |> min() == "2021-04-01"
+as.Date(paste(sociodemographic_characteristics$REF_DATE,"-01", sep="")) |> max() == "2023-07-01"
+
+sociodemographic_characteristics$VALUE |> min() >= 0
+
+distinct(sociodemographic_characteristics, GEO)
+
[…]

### 2024-04-22 paper/paper.qmd
Updated paper

@@
 #| message: false
-#| label: fig-trend4
+#| label: fig-gender1
 #| fig-cap: Perceived mental health fair or poor in Canada
@@
 #| message: false
-#| label: fig-trend2
+#| label: fig-gender2
 #| fig-cap: Perceived mental health fair or poor in Canada
@@
 #| message: false
-#| label: fig-trend4
+#| label: fig-age1
 #| fig-cap: Perceived mental health fair or poor in Canada
@@
 #| message: false
-#| label: fig-trend2
+#| label: fig-age2
 #| fig-cap: Perceived mental health fair or poor in Canada
@@
 #| message: false
-#| label: fig-trend2
+#| label: fig-other
 #| fig-cap: Perceived mental health fair or poor in Canada

### 2024-04-22 paper/paper.qmd
Removed subtitle

@@
 title: "Navigating Through Crisis: The Influence of COVID-19 on Canada’s Mental Health Problem"
-subtitle: "My subtitle if needed"
 author: 

### 2024-04-24 paper/paper.qmd
Updated

--- 2024-04-22 paper/paper.qmd
@@
 #| label: fig-trend1
-#| fig-cap: Perceived mental health fair or poor in Canada, 2002-2022
+#| fig-cap: Trends of Reported Fair or Poor Mental Health in Canada (2002-2022)
 
@@
 #| label: fig-trend2
-#| fig-cap: Perceived mental health fair or poor in Canada
+#| fig-cap: Trends of Reported Fair or Poor Mental Health in Canada (2021-2023)
 
@@
 #| label: fig-province1
-#| fig-cap: Perceived mental health fair or poor in Canada
+#| fig-cap: Provincial Comparison of Reported Fair or Poor Mental Health in Canada (2015-2022)
 
@@
 #| label: fig-province2
-#| fig-cap: Perceived mental health fair or poor in Canada
+#| fig-cap: Provincial Comparison of Reported Fair or Poor Mental Health in Canada (2021-2023)
 
@@
 #| message: false
-#| label: fig-trend4
-#| fig-cap: Perceived mental health fair or poor in Canada
+#| label: fig-gender1
+#| fig-cap: Gender Differences in Reported Fair or Poor Mental Health in Canada (2015-2022)
 
@@
 #| message: false
-#| label: fig-trend2
-#| fig-cap: Perceived mental health fair or poor in Canada
+#| label: fig-gender2
+#| fig-cap: Gender Differences in Reported Fair or Poor Mental Health in Canada (2021-2023)
 
@@
 #| message: false
-#| label: fig-trend4
-#| fig-cap: Perceived mental health fair or poor in Canada
+#| label: fig-age1
+#| fig-cap: Age-Related Trends in Reported Fair or Poor Mental Health in Canada (2015-2022)
 
@@
 #| message: false
-#| label: fig-trend2
-#| fig-cap: Perceived mental health fair or poor in Canada
+#| label: fig-age2
+#| fig-cap: Age-Related Trends in Reported Fair or Poor Mental Health in Canada (2021-2023)
 
@@
 #| message: false
-#| label: fig-trend2
-#| fig-cap: Perceived mental health fair or poor in Canada
+#| label: fig-other
+#| fig-cap: Mental Health Disparities Among Special Populations in Canada
 
@@
 ```
-
 # Discussion
 
-## First discussion point {#sec-first-point}
+## Evaluation of Research Goals
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+In this paper, we embarked on a nuanced exploration of mental health in Canada over two critical decades, employing datasets that encapsulate the pre and intra-pandemic eras.
+Through this retrospective lens, we've charted the evolution of public perceptions of mental health and the stark rise in the prevalence of fair or poor mental health reports, particularly post-2020.
+The data delineate a landscape where mental health has become a prominent concern, reflecting both the immediate shock of the pandemic and underlying systemic vulnerabilities.
 
-## Second discussion point
+## New Perspectives on Mental Health in Canada
 
-## Third discussion point
+Our findings illuminate the heightened mental health challenges faced by Canadians during the COVID-19 pandemic.
+A significant revelation is the disparate impact across demographics, with particular strains on LGBTQ2+ communities, women, and the youth.
+These insights underscore the pandemic's multifaceted toll beyond the immediate health implications, permeating social, economic, and personal domains.
 
-## Weaknesses and next steps
+## Intersectional Impact and Systemic Factors
 
-Weaknesses and next steps should also be included.
+The data also cast light on the intersectional nature of mental health challenges, revealing how gender, sexuality, age, and ethnicity intersect to influence mental well-being.
+For instance, the elevated rates of mental health concerns among LGBTQ2+ populations signal a distressing reality of social stressors compounded by a health crisis.
+This prompts a discussion on systemic issues like discrimination and access to care that may exacerbate such disparities.
+
+## Limitations and Considerations
+
+Notwithstanding these insights, the study has its constraints.
+The varied data collection methodologies over time introduce complexity in direct comparisons.
+Additionally, self-reported data on mental health may be influenced by evolving societal attitudes towards mental health, potentially impacting trend analysis.
+
+## Pathways for Future Research
+
+Going forward, a longitudinal approach to data collection would be invaluable in tracing the long-term impact of the pandemic on mental health.
+Moreover, in-depth qualitative research could enrich our understanding of the lived experiences behind the statistics.
+Policymaking should pivot on these findings to provide targeted mental health support where it is most needed.
+
+# Conclusion
+
+The conversation on mental health, while brought into sharp focus by the pandemic, is far from new.
+However, our study reaffirms that mental health is an enduring public health priority.
+As we move towards recovery, let this research serve as a clarion call to policymakers and healthcare providers to fortify mental health support systems, ensuring they are robust, inclusive, and responsive to the evolving needs of the Canadian populace.
 

### 2024-04-24 paper/paper.qmd
Updated

@@
 
-The COVID-19 pandemic has spotlighted mental health as a critical component of public health, magnifying existing challenges and uncovering gaps within healthcare systems globally.
-In Canada, the impact of the pandemic on mental health has been profound, exacerbating disparities and highlighting the urgency of effective mental health care.
-The Centre for Addiction and Mental Health (CAMH) has identified a significant increase in mental health problems among Canadians, noting that the pandemic has led to unprecedented levels of anxiety and depression (CAMH, 2020).
-Additionally, the Public Health Agency of Canada reports on the elevated symptoms of anxiety and depression during the pandemic, further underscoring the need for comprehensive data analysis on this issue (Public Health Agency of Canada, 2021).
+The COVID-19 pandemic did not only threaten people's physical health but also potentially harm their mental health.
+In Canada, the impact of the pandemic on mental health has been profound.
+The Centre for Addiction and Mental Health [@CAMH] has identified a significant increase in mental health problems among Canadians, noting that the pandemic has led to unprecedented levels of anxiety and depression .
+Additionally, @PHAC reports on the elevated symptoms of anxiety and depression during the pandemic, further underscoring the need for comprehensive data analysis on this issue .
 
@@
 
-The study addresses the gap in longitudinal mental health research, particularly in the analysis of sociodemographic influences during a major global crisis.
-Through the comparison of mental health indicators before and during the pandemic, this research provides insights into the shifts in mental health perceptions and the effectiveness of the responses to these changes.
-Initial findings suggest notable disparities in mental health impacts across different demographic groups, indicating the need for targeted interventions.
+Through a comparative analysis of mental health indicators before and during the pandemic, this research reveals significant insights into the mental health problems in Canada.
+The findings highlight a concerning rise in reported mental health concerns across all provinces, genders, and age groups, with a particularly sharp increase observed since the begin of the pandemic.
+Furthermore, the analysis reveals disparities in mental health experiences among specific populations, including women, youth, older adults, LGBTQ2+ individuals, and Indigenous communities.
+These findings underscore the need for targeted interventions and support systems that address the unique challenges faced by diverse groups within Canada.
 
-The significance of this research lies in its contribution to the evidence base informing mental health policy and practice.
-By identifying the demographic factors associated with varying mental health outcomes, the study aids policymakers and healthcare providers in developing tailored strategies to address the unique challenges faced by diverse groups within Canada.
-*TBD*
+This research contributes valuable evidence to inform mental health policy and practice in Canada.
+By identifying the demographic factors associated with varying mental health outcomes, the study aids policymakers and healthcare providers in developing tailored strategies and allocating resources effectively.
+The paper is structured as follows: Section 2 provides a detailed overview of the datasets and methodologies employed; Section 3 presents the key findings and trends; Section 4 delves into a discussion of the implications and connects the findings to broader contexts; and Section 5 concludes by summarizing the key insights and proposing future research directions.
 
@@
 ### Mental Health Indicators Dataset
+
 Data from 2002 and 2012 in the Mental Health Indicators Dataset were derived from national health databases, capturing key statistics on the prevalence of mental disorders and utilization of mental health services.
@@
 ### Perceived Mental Health by Age Group
+
 The Perceived Mental Health by Age Group dataset spans from 2015 to 2022 and utilizes a consistent methodological framework to assess self-reported mental health.
@@
 ### Perceived Mental Health by Gender and Sociodemographic Characteristics
+
 The data for 2021 to 2023 from this dataset is sourced from the Canadian Community Health Survey (CCHS), which meticulously gathers data on mental health perceptions influenced by gender and other sociodemographic factors such as income, education, and employment.
@@
 
-# Results
+# Results  {#sec-results}
 
@@
 The results section will further dissect these trends by province, gender, age, and within special populations to understand the nuances of this mental health challenge.
+
 ```{r}
@@
 ```
+
 ## By Province
@@
 The consistent rise across provinces post-2020, however, points to the overarching influence of the pandemic on mental health at a national level.
+
 ```{r}
@@
 These findings highlight the significance of gender-responsive mental health resources and interventions, particularly in addressing the unique challenges that have emerged during the pandemic.
+
 ```{r}
@@
 ```
+
 ## By Age
+
 The first graph spans from 2016 to 2022 and illustrates the percentage of people within various age brackets reporting their mental health as fair or poor.
@@
 Historical context, ongoing systemic inequalities, and intergenerational trauma are likely contributing factors and underscore the importance of providing culturally competent care and support.
+
 ```{r}
@@
 ```
-# Discussion
+
+# Discussion  {#sec-discussion}
 
@@
 
-# Conclusion
+# Conclusion {#sec-conclusion}
 

### 2024-04-25 paper/paper.qmd
Updated paper

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study analyzes the impact of the COVID-19 pandemic on mental health across various demographics in Canada, using data from 2002 to 2023 to assess trends and disparities.
+Our findings reveal a significant increase in reports of fair or poor mental health, particularly pronounced among young adults and vulnerable groups such as LGBTQ2+ and women.
+The research highlights the urgent need for improved mental health interventions, especially after the global health crises, which significantly affect certain populations.
+This paper contributes to our understanding of how pandemics worsen existing mental health vulnerabilities and shows the importance of responsive and inclusive health policy making."
 format: pdf
@@
 library(here)
-library(knitr)
+library(kableExtra)
 age <- read_parquet(file = here("data/analysis_data/age.parquet"))
@@
 
+The paper is organized into five main sections.
+@sec-data elaborates on the datasets and methodologies employed in this study.
+@sec-results details the statistical findings and examines trends by province, gender, age, and among other affected populations.
+@sec-discussion engages in a discussion of the broader implications of these findings, considers the study's limitations, and suggests avenues for future research.
+Finally, @sec-conclusion concludes the paper, summarizing key insights and emphasizing the necessity for informed policy-making in mental health.
+
+To analyze the data, I worked on R [@citeR], using tidyverse[@tidyverse], here [@here] and arrow [@arrow] packages.
+Tables in the report were created with kableExtra [@kableExtra] and figures were created with ggplot2 which was built in tidyverse [@tidyverse].
+
+The estimand is the difference in the probability of reporting mental health as fair or poor before versus during the COVID-19 pandemic.
+This analysis focuses on Canadian residents across various provinces and key demographic groups.
+
 # Data {#sec-data}
 
-This study uses four carefully curated datasets from Statistics Canada, each providing valuable insights into the mental health perceptions of Canadians.
-These datasets are pivotal for understanding the evolution of mental health across different demographics and geographical regions from 2002 to 2023.
+## Overview of Datasets
+This study uses four datasets from Statistics Canada, each providing valuable insights into the mental health perceptions of Canadians.
+These datasets are important for understanding the evolution of mental health across different demographics and geographical regions from 2002 to 2023.
 
-## Overview of Datasets
+The datasets contain some variables in common, including "REF_DATE" for time, "GEO" for location,  "Sex" or "Gender", alongside "Indicators" and "VALUE" which detail specific mental health metrics and their quantitative assessments.
+Apart from these, "Age" can be found in the Mental Health Indicators Dataset and The Perceived Mental Health by Age Group Dataset.
+The Perceived Mental Health by Gender and Sociodemographic Characteristics Dataset has a special variable "Sociodemographic_characteristics", which includes a range of detailed factors such as age, identities, race etc.
 
-The data for this research come from national surveys conducted by Statistics Canada, specifically designed to monitor various aspects of mental health across the country.
-Each dataset targets different demographic variables, such as age, gender, and province, and covers distinct periods critical for analyzing long-term trends and the immediate effects of the COVID-19 pandemic.
+### Mental Health Indicators Dataset (2002 and 2012)
 
-### Mental Health Indicators Dataset
-
-Data from 2002 and 2012 in the Mental Health Indicators Dataset were derived from national health databases, capturing key statistics on the prevalence of mental disorders and utilization of mental health services.
-The information is gathered through comprehensive health surveys and administrative records, providing robust measures of mental health status and service outcomes.
-
-@tbl-1 shows the first 5 rows in the dataset, REF_DATE stands for the time that the data was gathered, GEO denotes the geographical ,
+The dataset gathered by @mhi is important as it provides historical information of mental health status among Canadians.
+It offers valuable insights into the prevalence of various mental disorders and the utilization of mental health services across different time periods, helping to identify long-term trends in mental health care and outcomes.
 
 ```{r}
-#| label: tbl-1
+#| label: tbl-mhi
 #| tbl-cap: First 5 Rows in Mental Health Indicators Dataset 
@@
 
-kable(head(mental_health_indicators,5))
+kable(head(mental_health_indicators,5), "latex", booktabs = T)|>
+  kable_styling(latex_options = c("striped", "scale_down"))
+
 ```
 
-### Perceived Mental Health by Age Group
+### Perceived Mental Health by Age Group Dataset (2015-2022)
 
-The Perceived Mental Health by Age Group dataset spans from 2015 to 2022 and utilizes a consistent methodological framework to assess self-reported mental health.
-The data collection process involves standardized questionnaires distributed to a representative sample of Canadian adults across various age groups.
-This approach ensures that the data reflect a broad spectrum of the population, excluding those in full-time residence, on Indigenous reserves, or non-permanent residents, thus focusing on a generalizable sample of the Canadian populace.
+Focusing on self-reported mental health across different age brackets, this dataset enables an analysis of how mental health perceptions vary with age [@age].
+It covers more recent years and is instrumental in understanding the shifts in mental health attitudes and conditions in the context of modern societal changes, including the pre and post periods of the COVID-19 pandemic.
 
-### Perceived Mental Health by Gender and Sociodemographic Characteristics
+```{r}
+#| label: tbl-age
+#| tbl-cap: First 5 Rows in Perceived Mental Health by Age Group Dataset 
+#| echo: false
 
-The data for 2021 to 2023 from this dataset is sourced from the Canadian Community Health Survey (CCHS), which meticulously gathers data on mental health perceptions influenced by gender and other sociodemographic factors such as income, education, and employment.
-This survey selects participants through a stratified sampling design aimed at capturing diverse sociodemographic landscapes across Canada, enhancing the dataset's relevance for analyzing intersectional impacts on mental health.
+kable(head(age,5), "latex", booktabs = T)|>
+  kable_styling(latex_options = c("striped", "scale_down"))
+```
 
-### Detailed Measurement Discussion
+### Perceived Mental Health by Gender and Sociodemographic Characteristics Dataset (2021-2023)
 
-Both the CCHS and other Statistics Canada surveys employ rigorous data verification and validation processes.
-Quality control measures include logical checks during data collection and detailed follow-ups for non-responses.
-Post-stratification weighting adjusts for any discrepancies due to non-response rates, and seasonal adjustments are made to account for the timing of data collection.
-These methodologies ensure that the data are robust, reliable, and accurately reflect the Canadian population’s health status.
+This dataset [@sc] is essential for exploring how different sociodemographic factors such as gender, identities, and age impact mental health perceptions.
+Its recent collection provides current data, crucial for assessing the ongoing effects of the COVID-19 pandemic on various demographic groups within Canada.
 
-The CCHS's use of health-specific, standardized questionnaires is designed to capture nuanced information on health determinants and outcomes, which is critical for public health research and intervention planning.
-Conversely, Statistics Canada’s broader surveys provide comprehensive data across a wider range of topics, offering valuable insights into the interplay between health and other societal factors.
+```{r}
+#| label: tbl-sc
+#| tbl-cap: First 5 Rows in Perceived Mental Health by Gender and Sociodemographic Characteristics Dataset
+#| echo: false
 
-### Comparison of Data Collection Specifics
+kable(head(sc,5), "latex", booktabs = T)|>
+  kable_styling(latex_options = c("striped", "scale_down"))
+```
 
-CCHS is highly adaptable with its questionnaire content, allowing for the inclusion of regional health issues of specific interest, which is crucial for targeted public health interventions.
-This adaptability in data collection, combined with its detailed focus on health, makes CCHS uniquely suited for studies like ours that require specific health-related insights.
+### Perceived Mental Health by Province Dataset (2021-2023)
+Also sourced from the same survey as the Perceived Mental Health by Gender and Sociodemographic Characteristics Dataset[@province], it gathers data on mental health perceptions across different Canadian provinces, examining regional differences and location-specific factors that may influence mental health outcomes.
+```{r}
+#| label: tbl-province
+#| tbl-cap: First 5 Rows in Perceived Mental Health by Province Dataset
+#| echo: false
 
-In summary, the methodologies used by the Canadian Community Health Survey and other Statistics Canada surveys are both robust and tailored to meet different informational needs.
-Understanding these methodologies helps in accurately interpreting the data collected on mental health perceptions, ensuring that research findings are based on reliable and representative data.
+kable(head(province,5), "latex", booktabs = T)|>
+  kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+## Measurement
+
+### Measurement Techniques for the Age Dataset
+The "Perceived Mental Health by Age Group" dataset, derived from the survey documented at Statistics Canada's IMDB page [@cchs], using a comprehensive methodology to capture how perceptions of mental health vary across different age groups in Canada.
+This survey involved administering standardized questionnaires to a representative sample of Canadian adults.
+The process was designed to ensure that the collected data accurately reflect the mental health statuses across varying age demographics, which are critical in identifying trends and specific age-related challenges in mental health.
+The survey’s methodology, focusing on age as a key variable, allows researchers to analyze how mental health perceptions change with age, providing insights into the lifecycle dynamics of mental health issues.
+
+### Measurement Techniques for the Sociodemographic Characteristics and Province Datasets
+The datasets focusing on "Perceived Mental Health by Gender and Sociodemographic Characteristics" and "Perceived Mental Health by Province" were both derived from the survey detailed on Statistics Canada's IMDB [@css].
+This survey collected data across various provinces, incorporating a broad range of sociodemographic characteristics, including gender, income, education, and employment status, among others.
+The methodology employed in this survey involved both direct questionnaire techniques and sophisticated stratified sampling designs.
+These methods ensured that the data captured not only reflect the diverse sociodemographic landscape of Canada but also allow for an in-depth analysis of how different factors influence mental health perceptions regionally and across various sociodemographic groups.
+
+### Key Differences and Considerations
+
+The key differences between the Canadian Community Health Survey (CCHS) and the Canadian Social Survey (CSS) lie in their focus and methodologies.
+The CCHS focused on health-specific data, making it suitable for in-depth public health studies and policies.
+It used traditional personal and telephone interviews to gather detailed health information.
+In contrast, the CSS covers broader social and economic topics, using both online and traditional methods to capture a wider demographic.
+This makes the CSS more adaptable to examining the societal factors influencing overall well-being.
+These distinctions are crucial for choosing the right dataset for specific research needs, with the CCHS offering depth in health topics and the CSS providing a broader view of societal well-being.
 
@@
 Prior to examining the trends, it's crucial to note that the datasets used for this analysis were collected using different methodologies.
-Annual data up to 2020 required a distinct approach from the quarterly data collected during the pandemic's peak.
-To preserve the accuracy of the analysis, the results are presented in separate graphs corresponding to these differing collection frequencies.
+To preserve the accuracy of the analysis, the results are presented in separate graphs.
 
@@
 
-The results from the analyzed datasets reveal significant trends in the mental health landscape across Canada.
+The results from the analyzed datasets reveal significant trends in mental health problems across Canada.
 Initially, the data show a gentle increase in reported mental health concerns from the early 2000s up until 2010.
 Post-2010, there is a noticeable dip, followed by a steady climb in mental health issues leading up to 2015.
-However, the most striking feature of this trend is the sharp and substantial escalation in mental health concerns starting from around 2020, which corresponds with the onset of the COVID-19 pandemic.
+However, the most striking feature of this trend is the sharp increase in mental health concerns starting from around 2020, which corresponds with the onset of the COVID-19 pandemic.
 
 @fig-trend1 plots the percentage of individuals reporting mental health concerns from 2005 to the present.
-From 2005 to around 2015, the trajectory exhibits minor fluctuations with a gradual increase.
+From 2005 to around 2015, there were minor fluctuations with a gradual increase.
 However, in the years following 2015, there is a distinct upward surge, with the percentages almost doubling by 2020 and continuing to rise steeply.
-This dramatic increase highlights the profound impact that the pandemic has had on mental health, as isolation, stress, economic uncertainty, and health-related fears took a toll on the population.
+This dramatic increase highlights the impact that the pandemic has had on mental health, as isolation, stress, economic uncertainty, and health-related fears affect the population.
 
-@fig-trend2 provides a more granular view of the recent period, showing quarterly data from 2021 through to mid-2023.
+@fig-trend2 provides a view of the recent period, showing quarterly data from 2021 through to mid-2023.
 This period captures the ongoing effects of the pandemic.
@@
 
-Together, these graphs paint a picture of a population that experienced a significant increase in mental health concerns during a global crisis.
+Together, these graphs show that Canadians are experiencing a significant increase in mental health concerns during a global crisis.
 The data underscore the need for ongoing monitoring and support to address mental health at a community and national level.
-The results section will further dissect these trends by province, gender, age, and within special populations to understand the nuances of this mental health challenge.
 
@@
 
-The results underscore the regional variations in mental health across Canadian provinces.
-In the graph illustrating data from 2016 to 2022, we see a common upward trend among all provinces, with a particularly steep increase following 2020, which aligns with the global onset of the COVID-19 pandemic.
-Prior to 2020, there were variations in the VALUE among provinces, but the gap between regions with the highest and lowest reported VALUE narrows as all regions trend upwards significantly in the subsequent years.
+This part demonstrates the regional variations in mental health across Canadian provinces.
+In @fig-province1, we see a common upward trend among all provinces, with a particularly steep increase following 2020, which aligns with the global onset of the COVID-19 pandemic.
+Prior to 2020, there were variations in the population among provinces, but the gap between regions with the highest and lowest reported values narrowed as all regions trended upwards significantly in the following years.
 
-Alberta, British Columbia, and Ontario exhibit some of the highest increases, suggesting a pronounced impact of the pandemic and its associated factors on mental health in these regions.
-However, it is notable that the trend is national, as even provinces like Prince Edward Island and Newfoundland and Labrador, which initially had some of the lowest reported VALUE, also show a sharp rise.
+Alberta, British Columbia, and Ontario has some of the highest increases, showing an obvious impact of the pandemic and its associated factors on mental health in these regions.
+However, it is notable that the trend is national, as even provinces like Prince Edward Island and Newfoundland and Labrador, which initially had some of the lowest reported values, also show a sharp rise.
 
-The second graph, representing quarterly data from 2021 to 2023, provides a more nuanced picture with noticeable fluctuations within each province.
+@fig-province2 represents data from 2021 to 2023, shows the fluctuations within each province.
 These variations could be indicative of the changing conditions during the pandemic, such as waves of COVID-19 cases, differing provincial public health restrictions, and economic instability.
@@
 Both graphs indicate that while the trend in mental health concerns is increasing across Canada, there are provincial specificities that suggest the impact of local factors.
-These could include economic conditions, the accessibility of mental health services, public health policies, and the social resilience of communities.
-The consistent rise across provinces post-2020, however, points to the overarching influence of the pandemic on mental health at a national level.
+These could include economic conditions, the accessibility of mental health services, public health policies, and support of communities.
+The consistent rise across provinces post-2020, however, points to the influence of the pandemic on mental health at a national level.
 
@@
   geom_line()+
+  labs(x = "Year", y = "Percentage") +
   theme_linedraw()
@@
 ggplot(temp, aes(x= REF_DATE, y= VALUE, group = GEO, color = GEO))+
+  scale_x_discrete(guide = guide_axis(angle = 90)) +
   geom_line()+
+  labs(x = "Year", y = "Percentage") +
   theme_linedraw()
@@
 
-The gender-based analysis reveals telling trends in the self-reported mental health status among Canadians, distinguishing between those who consider their mental health to be fair or poor.
+In examining the impact of the COVID-19 pandemic on mental health by gender, our analysis reveals distinct trends that highlight considerable differences between males and females in reporting fair or poor mental health.
 
-From 2016 to 2022, the first graph shows an upward trajectory for both males and females, with a notably sharper rise for females.
-By 2022, the percentage of females reporting their mental health as fair or poor substantially exceeds that of males, suggesting gender-specific impacts.
-The increase that coincides with the pandemic period may reflect the compounded stressors that women have faced, including potential increases in caregiving responsibilities, economic pressures, or other psychosocial factors exacerbated by the pandemic.
+From @fig-gender1, there is a clear rise in mental health concerns for both genders, with females consistently reporting higher levels of fair or poor mental health compared to males.
+This difference became more significant during the pandemic, suggesting that the crisis may have increased existing gender-related mental health vulnerabilities.
 
-In the quarterly data from 2021 to mid-2023, depicted in the second graph, the fluctuations are more pronounced, with both genders experiencing variability in their mental health reporting.
-The trend for women is characterized by higher peaks, indicating more pronounced shifts in mental health status as compared to men.
-This volatility could be influenced by the ebb and flow of pandemic-related stresses and public health measures.
+@fig-gender2 offers a closer look at these trends, showing significant fluctuations that correspond with ongoing pandemic developments and public health measures.
+These periodic changes are especially notable among females, who show greater peaks and troughs in mental health reporting compared to males.
+This suggests that women's mental health may be more sensitive to the immediate impacts of COVID-related stressors.
 
-For both sexes combined, the overall trend is less erratic, pointing to the importance of disaggregating data by gender to fully capture the differential impacts on mental health.
-These findings highlight the significance of gender-responsive mental health resources and interventions, particularly in addressing the unique challenges that have emerged during the pandemic.
-
+This gender-specific analysis not only shows the differential impact of the pandemic on mental health between males and females but also highlights the necessity for targeted mental health interventions that address the unique challenges faced by women during such crises.
 ```{r}
@@
   geom_point()+
+  labs(x = "Year", y = "Percentage") +
   scale_color_manual(values = colors) +
@@
   labs(x = "Year", y = "Percentage") +
+  scale_x_discrete(guide = guide_axis(angle = 90)) +
   scale_color_manual(values = colors) +
@@
 
-The first graph spans from 2016 to 2022 and illustrates the percentage of people within various age brackets reporting their mental health as fair or poor.
-A distinct trend is visible across all age groups, showing an increase in self-reported fair or poor mental health, with a marked acceleration from 2020 onwards, likely due to the COVID-19 pandemic.
-The oldest age group, those aged '65 years and over,' displays the most pronounced rise, indicating this demographic may have been particularly impacted during these years.
-Meanwhile, younger age groups also show a worrying increase, albeit less steep, suggesting widespread implications across generations.
+The analysis of mental health trends by age reveals significant variations in how different age groups have perceived their mental health from 2016 to 2022, with notable impacts during the COVID-19 pandemic.
 
-The second graph, which provides quarterly data from early 2021 to mid-2023, presents a more granular view, capturing the fluctuations within and between age groups during the pandemic.
-Notably, the '15 to 24 years' and '25 to 34 years' categories exhibit significant volatility, with peaks and valleys that could correlate with periods of pandemic waves, social restrictions, and economic impacts.
-The '65 years and over' group continues to report higher rates of fair or poor mental health, although with less fluctuation, possibly reflecting ongoing concerns about health risks and social isolation among seniors.
+In @fig-age1, an increasing trend in reporting fair or poor mental health was observed across all age groups, which became significantly worse due to the breakout of the pandemic.
+However, the patterns varied markedly between age cohorts.
+The data show that those aged 18 to 34 years reported the most substantial increases in mental health concerns.
+This suggests that the pandemic's effects have been particularly profound on the mental health of the young people in society, possibly due to different stressors relevant to their stages in life such as educational disruptions and economic downturns.
 
-These age-focused graphs suggest that mental health challenges have been a critical concern for all age groups, with particular implications for the most vulnerable populations.
-The rise in reported mental health issues among the youth and elderly underscores the need for age-specific mental health services and support systems tailored to their unique circumstances and stressors.
-The data call for continued monitoring and interventions to address the mental health repercussions of the pandemic across all age demographics.
+@fig-age2 provides further insights into these trends, revealing fluctuations that align with pandemic waves and associated public health responses.
+Young adults demonstrated considerable fluctuations in their reported mental health, with young adults experiencing mental health challenges including social restrictions and periods of uncertainty.
+These reflect the particular vulnerability of younger individuals to the social and educational disruptions caused by the pandemic.
+
+This age-specific analysis showed the impact of the pandemic on mental health across the different age groups.
+It emphasizes that young adults need more mental health support.
+These findings are critical for developing strategic public health responses that effectively address the mental health fallout from the pandemic and support the resilience of all age groups.
 
@@
   geom_point()+
+  labs(x = "Year", y = "Percentage") +
   theme_linedraw()
@@
   labs(x = "Year", y = "Percentage") +
+  scale_x_discrete(guide = guide_axis(angle = 90)) +
   theme_linedraw()
@@
 The examination of mental health within specific demographic segments reveals significant disparities.
-The most affected population, with 48.4%, are LGBTQ2+ women, followed by the total population of LGBTQ2+ individuals at 44.5%.
+@tbl-other showed that the most affected population, with 48.4% reporting fair or poor mental health, are LGBTQ2+ women, followed by the total population of LGBTQ2+ individuals at 44.5%.
 LGBTQ2+ men report a lower percentage at 39.0%, indicating a gender-related variance within the LGBTQ2+ community.
@@
 #| message: false
-#| label: fig-other
-#| fig-cap: Mental Health Disparities Among Special Populations in Canada
+#| label: tbl-other
+#| tbl-cap: Mental Health Disparities Among Special Populations in Canada
 
@@
 temp2 <- temp1[order(-temp1$Average),]
-kable(head(temp2,15))
+kable(head(temp2,15), "latex", booktabs = T)|>
+  kable_styling(latex_options = c("striped", "scale_down"))
 
@@
 
-In this paper, we embarked on a nuanced exploration of mental health in Canada over two critical decades, employing datasets that encapsulate the pre and intra-pandemic eras.
-Through this retrospective lens, we've charted the evolution of public perceptions of mental health and the stark rise in the prevalence of fair or poor mental health reports, particularly post-2020.
-The data delineate a landscape where mental health has become a prominent concern, reflecting both the immediate shock of the pandemic and underlying systemic vulnerabilities.
+In this paper, we explored the mental health problems in Canada over two critical decades, employing datasets that encapsulate the pre and intra-pandemic eras.
+Through the analysis, we've shown the evolution of the public's mental health status.
+The data demonstrated that mental health has become an influential concern, reflecting both the shock of the pandemic and underlying systemic vulnerabilities.
 
@@
 
-Our findings illuminate the heightened mental health challenges faced by Canadians during the COVID-19 pandemic.
-A significant revelation is the disparate impact across demographics, with particular strains on LGBTQ2+ communities, women, and the youth.
-These insights underscore the pandemic's multifaceted toll beyond the immediate health implications, permeating social, economic, and personal domains.
+Our findings reviewed the mental health challenges faced by Canadians before and during the COVID-19 pandemic.
+A significant find is the impact across demographics, with particular strains on LGBTQ2+ communities, women, and the youth.
+These insights underscore the pandemic's influence beyond the immediate health implications, permeating social, economic, and personal domains.
 
@@
 
-The data also cast light on the intersectional nature of mental health challenges, revealing how gender, sexuality, age, and ethnicity intersect to influence mental well-being.
-For instance, the elevated rates of mental health concerns among LGBTQ2+ populations signal a distressing reality of social stressors compounded by a health crisis.
-This prompts a discussion on systemic issues like discrimination and access to care that may exacerbate such disparities.
+The data also discussed the intersectional nature of mental health challenges, revealing how gender, sexuality, age, and ethnicity intersect to influence mental well-being.
+For instance, the increasing rates of mental health concerns among LGBTQ2+ populations stated the reality of social stressors.
+This leads to a discussion on systemic issues like discrimination and access to care.
 
@@
 
-Notwithstanding these insights, the study has its constraints.
+Despite these insights, the study has its constraints.
 The varied data collection methodologies over time introduce complexity in direct comparisons.
-Additionally, self-reported data on mental health may be influenced by evolving societal attitudes towards mental health, potentially impacting trend analysis.
+Additionally, self-reported data on mental health may be influenced by evolving societal attitudes toward mental health, potentially impacting trend analysis.
 
-## Pathways for Future Research
+## Future Research
 
-Going forward, a longitudinal approach to data collection would be invaluable in tracing the long-term impact of the pandemic on mental health.
-Moreover, in-depth qualitative research could enrich our understanding of the lived experiences behind the statistics.
-Policymaking should pivot on these findings to provide targeted mental health support where it is most needed.
+For future research, to comprehensively assess the lasting impact of the COVID-19 pandemic on mental health, researchers could use longitudinal studies.
+These studies would track individuals over extended periods, providing insights into the long-term effects of the pandemic on various demographic groups.
+
+Another important area for future research is the evaluation of mental health policies implemented.
+Analyzing the effectiveness of these policies in real time could provide valuable feedback for policymakers, allowing for adjustments and more targeted interventions.
+
+Finally, doing global comparative studies may make the results more generalized.
+Understanding global best practices and learning from the successes and challenges faced by other nations could guide future strategies and interventions.
 
@@
 
-The conversation on mental health, while brought into sharp focus by the pandemic, is far from new.
-However, our study reaffirms that mental health is an enduring public health priority.
-As we move towards recovery, let this research serve as a clarion call to policymakers and healthcare providers to fortify mental health support systems, ensuring they are robust, inclusive, and responsive to the evolving needs of the Canadian populace.
+In conclusion, this study has provided an examination of the mental health problem in Canada during the period of the COVID-19 pandemic.
+By analyzing a range of datasets from 2002 to 2023, the results have highlighted significant increases in mental health concerns across all demographics with notable disparities.
+
+The findings underscore the profound impact of the pandemic on mental health, revealing an urgent need for better support systems to address the unique challenges faced by different populations, including young adults and vulnerable groups such as LGBTQ2+ individuals and women.
+These insights can be helpful for policymakers and healthcare providers, emphasizing the necessity to adapt and strengthen mental health services to cope with current and future challenges.
+
+Moving forward, the government should continue to monitor mental health trends to respond to emerging needs.
+Implementing the experiences learned from this research can lead to more helpful health systems that not only manage crises effectively but also support the ongoing well-being of the population.
+By doing so, we can ensure that mental health care in Canada remains responsive, inclusive, and proactive, reflecting the complexities of human health in a changing world.
 

### 2024-04-25 other/datasheet/datasheet.qmd
Finished the datasheet

---
title: "Datasheet for 'Perceived mental health, by gender and other selected sociodemographic characteristics'"
author: 
  - Zijun Meng
thanks: "Code and data are available at: https://github.com/FrankMengZJ/canada_mental_health"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This is a datasheet for 'Perceived mental health, by gender and other selected sociodemographic characteristics'. By providing information about its motivation, composition, collection process, preprocessing/cleaning/labeling, uses, distribution, and maintenance, the datasheet can help users have a better understanding of this dataset."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset [@sc] was created to fill a gap in understanding the mental health problems in Canada.They are structured to provide insights that inform policy-making, academic research, and public understanding.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Statistics Canada created the datasets. 
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - They were funded by the Canadian government.
4. *Any other comments?*
    - No.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances in the dataset primarily represent individual responses from Canadian residents aged 15 and over, living in private households across the 10 provinces. These responses capture a variety of social and economic aspects.
2. *How many instances are there in total (of each type, if appropriate)?*
	- The exact number of instances varies from cycle to cycle, but each instance represents a survey response from an individual participant. The number typically extends into the tens of thousands to ensure a representative sample of the population.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- It's a sample of the larger Canadian population, specifically designed to be representative geographically and demographically. Statistics Canada uses stratified sampling to ensure that all provinces and key demographic groups are proportionately represented. The representativeness is validated through statistical techniques and comparison with known population benchmarks.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each survey response (instance) in the dataset consists of both raw and processed data, encompassing answers to survey questions on topics ranging from health to social participation. This data is  processed to ensure accuracy and reliability before being made available for analysis.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Each instance includes labels related to the survey questions, such as health status or social activity levels, which can be used for analytical purposes to assess trends and correlations within the population.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	-  Yes. Statistics Canada employs various methods to handle missing data, including statistical imputation where appropriate [@css].
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	-  The dataset includes metadata that allows researchers to examine relationships, such as familial or household connections, when this information is relevant to the survey's focus.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Despite Statistics Canada used several ways to minimize the errors, considering it is a large scale survey,there might be some errors.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is self-contained and does not rely on external resources that could change over time, ensuring its stability and reliability for longitudinal studies.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- Data confidentiality is strictly maintained, with all personal identifiers removed to ensure that individual respondents cannot be directly or indirectly identified.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- The dataset does not contain offensive content as it is purely factual and focused on socio-economic factors.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- The dataset identifies sub-populations by demographic characteristics such as age and gender, which are crucial for analyzing trends across different segments of the population.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	-  It is not possible to identify individuals directly from the dataset due to the anonymization and aggregation of data prior to public release.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- While the dataset may contain sensitive information, such as health status or personal economic conditions, all data is anonymized and presented in aggregate form to protect individual privacy.
16. *Any other comments?*
	-  No.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was primarily acquired through direct survey responses reported by subjects, covering various socio-economic aspects. Validation of this data includes consistency checks and comparison against other national data sources to ensure reliability.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Data was collected using a combination of computer-assisted telephone interviewing (CATI) and online self-administered questionnaires. These methods are validated through pilot tests and methodological research to ensure effectiveness and accuracy.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset uses a probabilistic sampling design, stratifying by geographical and demographic characteristics to ensure a representative sample of the Canadian population.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Data collection was conducted by trained interviewers employed by Statistics Canada, compensated as per government employment standards.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data is collected quarterly to ensure timely updates on social trends. This timeframe matches the creation timeframe of the data, as each data release corresponds directly to its collection period.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	-As a government survey, the CSS follows strict ethical guidelines reviewed by Statistics Canada's ethical review board, ensuring all practices comply with national standards for research ethics.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Data was collected directly from individuals through structured survey methodologies, without involvement from third parties.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	-  Individuals were notified about the data collection at the beginning of each survey through a standard script which explains the purpose of the survey and how the data will be used.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Consent to collect data was obtained verbally at the beginning of each survey session, with interviewers explaining the use of the data and participants agreeing to proceed.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Participants can request the removal of their data or decline to participate at any stage of the survey, with mechanisms in place to ensure their data is not used in subsequent analyses if consent is withdrawn.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- An impact analysis is typically conducted to assess the potential effects of the dataset on subjects, ensuring no adverse consequences arise from its use. This analysis is part of Statistics Canada's standard practice for all its surveys.
12. *Any other comments?*
	- No.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- The data collected through the CSS used several preprocessing and cleaning steps to ensure accuracy and usability. This typically includes handling missing values, correcting data entry errors, and standardizing responses to ensure consistency across the dataset. Additionally, responses might be categorized or bucketed to facilitate analysis, particularly when dealing with ordinal or categorical data.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Statistics Canada typically retains both the "raw" survey data and the processed versions. The raw data, which includes all original responses with minimal modification, is often used for methodological testing or to support further research that may require re-analysis with different parameters.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- The specific software used for data preprocessing, cleaning, and labeling by Statistics Canada is generally not publicly available. 
4. *Any other comments?*
	- No.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- CSS has been used for various research tasks, primarily focusing on understanding social trends and impacts of events like the COVID-19 pandemic on Canadian society. 
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No.
3. *What (other) tasks could the dataset be used for?*
	- The CSS could be employed in predictive modeling for social policy outcomes, sociological research to track changes in cultural norms, and more extensive public health research.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- The way the CSS is collected and processed ensures a high level of data integrity and representativeness, which is crucial for equitable analysis. However, users must be cautious about interpreting the data without considering the context of the questions and the manner in which responses were gathered to avoid biases that could lead to unfair treatment or stereotyping of individuals or groups. 
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used for purposes that require medical or highly sensitive personal data, as it does not contain detailed personal health information or other sensitive personal details that are protected under privacy laws.
6. *Any other comments?*
	- No.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- No.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- On Statistics Canada. Yes, it's https://doi.org/10.25318/4510008001-eng.
3. *When will the dataset be distributed?*
	- February 13, 2024
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- No.  
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No.
7. *Any other comments?*
	- No.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Statistics Canada.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- https://www.statcan.gc.ca/en/reference/refcentre/index provides many ways to contact
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- The dataset is updated regularly, in line with the survey's quarterly cycle. Statistics Canada is responsible for these updates and communicates them through its official website and via release announcements that can be subscribed to via email or RSS feeds.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- For datasets involving personal data, Statistics Canada adheres to strict privacy laws which dictate data retention policies. Participants are generally informed about the retention period during the survey consent process. This period aligns with the purpose of the data collection and is enforced through internal data governance policies.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- No. On their official website.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- No.
8. *Any other comments?*
	- No.

\newpage

# References

### 2024-04-25 paper/paper.qmd
Fine tuning

@@
 
-The COVID-19 pandemic did not only threaten people's physical health but also potentially harm their mental health.
+The COVID-19 pandemic has not only threatened people's physical health but also potentially harmed their mental health.
 In Canada, the impact of the pandemic on mental health has been profound.
@@
 
-The datasets contain some variables in common, including "REF_DATE" for time, "GEO" for location,  "Sex" or "Gender", alongside "Indicators" and "VALUE" which detail specific mental health metrics and their quantitative assessments.
+The datasets share several common variables, including 'REF_DATE' for time, 'GEO' for location, and 'Sex' or 'Gender', along with 'Indicators' and 'VALUE', which detail specific mental health metrics and their quantitative assessments.
 Apart from these, "Age" can be found in the Mental Health Indicators Dataset and The Perceived Mental Health by Age Group Dataset.
@@
 
-The results from the analyzed datasets reveal significant trends in mental health problems across Canada.
+The analysis of these datasets reveals significant trends in mental health issues across Canada.
 Initially, the data show a gentle increase in reported mental health concerns from the early 2000s up until 2010.
@@
 
-The data also discussed the intersectional nature of mental health challenges, revealing how gender, sexuality, age, and ethnicity intersect to influence mental well-being.
+The data also discuss the intersectional nature of mental health challenges, revealing how gender, sexuality, age, and ethnicity intersect to influence mental well-being.
 For instance, the increasing rates of mental health concerns among LGBTQ2+ populations stated the reality of social stressors.
@@
 
-Despite these insights, the study has its constraints.
-The varied data collection methodologies over time introduce complexity in direct comparisons.
-Additionally, self-reported data on mental health may be influenced by evolving societal attitudes toward mental health, potentially impacting trend analysis.
+Despite these insights, the study faces certain constraints.
+Varied data collection methodologies over time introduce complexities in making direct comparisons.
+Moreover, self-reported data on mental health may be influenced by evolving societal attitudes, which could potentially impact trend analyses.
+Furthermore, the quantitative focus might not fully capture the complexities of individuals' mental health experiences, underscoring the need for qualitative approaches to deepen our understanding.
+These limitations indicate that future research should incorporate methodological enhancements to ensure a more comprehensive analysis.
 
@@
 
-For future research, to comprehensively assess the lasting impact of the COVID-19 pandemic on mental health, researchers could use longitudinal studies.
+For future research, researchers could use longitudinal studies to comprehensively assess the lasting impact of the COVID-19 pandemic on mental health.
 These studies would track individuals over extended periods, providing insights into the long-term effects of the pandemic on various demographic groups.

### 2024-04-25 scripts/01-download_data.R
Updated preamble

@@ -1,26 +1,15 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data
+# Author: Zijun Meng
+# Date: 24 April 2024
+# Contact: zijun.meng@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
-#### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+# Download data from https://open.canada.ca/data/en/dataset/a9863f45-6a1a-4277-ae0f-2d9cb61d413a, https://doi.org/10.25318/4510008001-eng, 
+# https://doi.org/10.25318/4510007901-eng and https://doi.org/10.25318/1310009601-eng, theb put the files in "data/raw_data"
 
-
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
 
          

### 2024-04-25 scripts/02-data_cleaning.R

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw data in the 4 datasets.
+# Author: Zijun Meng
+# Date: 24 April 2024
+# Contact: zijun.meng@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Download the datasets and put them in the designated location.
 
 #### Workspace setup ####
 library(tidyverse)
@@ -26,6 +25,7 @@ mental_health_indicators <-
   filter(!is.na(VALUE)) |>
   select(REF_DATE, GEO, Age, Sex, Indicators, VALUE)
 
+
 # Clean 'Perceived mental health, by age group' dataset
 age <-
   age |>
@@ -33,8 +33,8 @@ age <-
   rename(Age='Age group') |>
   select(REF_DATE, GEO, Age, Sex, Indicators, VALUE)
 
-# Clean 'Perceived mental health, by gender and other selected sociodemographic characteristics' dataset
 
+# Clean 'Perceived mental health, by gender and other selected sociodemographic characteristics' dataset
 province <-
   province |>
   filter(Statistics == 'Percentage of persons') |>

### 2024-04-25 scripts/03-test_data.R

@@ -1,12 +1,10 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests the 4 datasets.
+# Author: Zijun Meng
+# Date: 24 April 2024
+# Contact: zijun.meng@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: Run 02-data_cleaning.R
 
 #### Workspace setup ####
 library(tidyverse)
